---
title: "Reproducibility"
author: "Mackenzie Gill"
date: "10/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Data
Importing data files into an inputs directory:

```{r}
setwd("~/OneDrive - Colostate/570/AREC570/inputs") #setting working directory
library(readr)
gdp_52_67 <- read_csv("gdp_52-67.csv") #GDP from 1952-1967
le_52_67 <- read_csv("le_52-67.csv") #Life expectancy from 1952-1967


```

## creating a single table of variables

```{r}

t<- merge(gdp_52_67, le_52_67 , by= c("continent","country", "year"))
t

```
## Canada 
The first graph shows that life expectancy has been increasing over this period of time in Canada. The second graph shows that per capita GDP is also increasing over this period of time in Canada. The simialr trend observed suggests that these factors may be correlated. 


```{r}
Canada<-subset(t, country=="Canada")
Canada

gdpPercap<- (Canada$gdp / Canada$pop)
gdpPercap

GdpPercap<- plot(Canada$year, gdpPercap,
      main = "GDP per Capita from 1952-1967", 
      ylab= "GDP per capita", xlab="Year", type = "l", col= "blue") 

lifeExp<- plot(Canada$year, Canada$lifeExp,
               main= "Life Expectancy from 1952-1967", 
               ylab = "Life expectancy", xlab = "Year", type = "l", col="red")

```
## Life Expectancy and GDP 
So now, we can see if GDP per capita leads to higher life expectancy. We expect the relationship between GDP per capita and life expectancy to be positive. To test this, we regress life expectancy from 140 countries on GDP per-capita. The results are displayed in Table 1.

```{r}
#install.packages("sjPlot")
#install.packages("sjmisc")
#install.packages("sjlabelled")

library(sjPlot)
library(sjmisc)
library(sjlabelled)


gdpPercap2<- (t$gdp / t$pop)
lifeexp<- (t$lifeExp)

M<- lm(lifeexp~gdpPercap2, data = t)
summary(M)

tab_model(M, collapse.ci = TRUE, p.style = "stars", digits = 4, title = "Table 1")

```

The estimated coefficient on gdpPercap is positive and statistically significant at α = 0.001. This implies that for a $10,000 increase in GDP per capita, life expectancy increases by five years.
This linear model implies that by increasing GDP, we can increase life expectancy continuously.
This is not realistic, as there is a limit to humans' potential life expectancy. Thus, further research should focus on the nonlinear relationship between these factors.




