---
title: "Reproducibility"
author: "Mackenzie Gill"
date: "10/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Data
Importing data files into an inputs directory:

```{r}
setwd("~/OneDrive - Colostate/570/AREC570/inputs")
library(readr)
gdp_52_67 <- read_csv("gdp_52-67.csv") #GDP from 1952-1967
le_52_67 <- read_csv("le_52-67.csv") #Life expectancy from 1952-1967


```

## creating a single table

```{r}

t<- merge(gdp_52_67, le_52_67 , by= c("continent","country", "year"))
t

```
## Canada 
Figure 1 shows that life expectancy has been increasing over time in Canada. That is good. Figure 2 shows that Canad’s GDP per capita is also increasing over time. Go Canada!
I wonder if these two data are correlated?

```{r}
Canada<-subset(t, country=="Canada")
Canada

gdpPercap<- (Canada$gdp / Canada$pop)
gdpPercap

GdpPercap<- plot(Canada$year, gdpPercap,
      main = "GDP per Capita from 1952-1967", 
      ylab= "GDP per capita", xlab="Year", type = "l", col= "blue") 

lifeExp<- plot(Canada$year, Canada$lifeExp,
               main= "Life Expectancy from 1952-1967", 
               ylab = "Life expectancy", xlab = "Year", type = "l", col="red")

```
## Life Expectancy and GDP 
The objective of this analysis is to test whether GDP per capita leads to higher life expectancy. We hypothesize that the relationship between GDP per capita and life expectancy is positive. We regress life expectancy from 140 countries on GDP per-capita to test our hypothesis. The results are displayed in Table 1.

```{r}
#install.packages("sjPlot")
#install.packages("sjmisc")
#install.packages("sjlabelled")


library(sjPlot)
library(sjmisc)
library(sjlabelled)




gdpPercap2<- (t$gdp / t$pop)
lifeexp<- (t$lifeExp)

M<- lm(lifeexp~gdpPercap2, data = t)
summary(M)

tab_model(M, collapse.ci = TRUE, p.style = "stars", digits = 4)

```

The coefficient on gdpPercap is positive and statistically significant at α = 0.05, which indicates that for every $10,000 increase in GDP per capita, life expectancy increases by five years.
We can conclude that by continually increasing GDP we can increase life expectancy.
Further research should focus on the nonlinear relationship between life expectancy and GDP per capita.




